Vesiculation pathways in clathrin-mediated endocytosis

Version 1

Abstract

Abstract

During clathrin-mediated endocytosis, a patch of flat plasma membrane is internalized to form a vesicle. In mammalian cells, how the clathrin coat deforms the membrane into a vesicle remains unclear and two main hypotheses have been debated. The “constant area” hypothesis assumes that clathrin molecules initially form a flat lattice on the membrane and deform the membrane by changing its intrinsic curvature while keeping the coating area constant. The alternative “constant curvature” hypothesis assumes that the intrinsic curvature of the clathrin lattice remains constant during the formation of a vesicle while the surface area it covers increases. Previous experimental studies were unable to unambiguously determine which hypothesis is correct. In this paper, we show that these two hypotheses are only two extreme cases of a continuum of vesiculation pathways if we account for the free energies associated with clathrin assembly and curvature generation. By tracing the negative gradient of the free energy, we define vesiculation pathways in the phase space of the coating area and the intrinsic curvature of clathrin coat. Our results show that, overall, the differences in measurable membrane morphology between the different models are not as big as expected, and the main differences are most salient at the early stage of endocytosis. Furthermore, the best fitting pathway to experimental data is not compatible with the constant-curvature model and resembles a constant-area-like pathway where the coating area initially expands with minor changes in the intrinsic curvature, later followed by a dramatic increase in the intrinsic curvature and minor change in the coating area. Our results also suggest that experimental measurement of the tip radius and the projected area of the clathrin coat will be the key to distinguish between models.

Introduction

Clathrin-mediated endocytosis (CME) is a fundamental cellular process to transport lipids, membrane proteins and extracellular cargo molecules into the cell (McMahon and Boucrot, 2011; Sorkin and Puthenveedu, 2013; Lu et al., 2016; Kaksonen and Roux, 2018; Lacy et al., 2018; Mettlen et al., 2018). In mammalian cells, a small patch of flat plasma membrane is shaped into a spherical vesicle when CME occurs (Avinoam et al., 2015). Clathrin molecules are essential for the membrane remodelling process. They are made of three subunits that form a triskelion, which further assemble into a cage-like structure in vitro (Musacchio et al., 1999; Shraiman, 1997). The minimum cages contain 16 polygons (Fotin et al., 2004) and the most commonly observed ones are semi-regular icosahedral cages (Cheng et al., 2007; Dannhauser and Ungewickell, 2012; Heuser, 1980). Two main hypotheses are under debate regarding how the clathrin coat scaffolds the flat membrane into a spherical vesicle in vivo (Avinoam et al., 2015; Chen and Schmid, 2020; Frey and Schwarz, 2020; Kaksonen and Roux, 2018; Scott et al., 2018). The constant area model asserts that the clathrin molecules initially polymerize into a flat lattice with a regularly arranged hexagonal structure, and later reorganization of the bonds between adjacent clathrins results in formation of pentagons in the hexagonal lattice, which in turn leads to curvature generation of the clahtrin coat (Fotin et al., 2004; Heuser and Anderson, 1989) (Figure 1a). Adhesion of clathrin molecules with the substrate has been suggested to contribute a flattening force that prevents curvature generation. Release of the flattening force therefore could induce curvature of the clathrin coat with preloaded pentagons (Sochacki et al., 2021). The alternative constant curvature model asserts that the intrinsic curvature of the elements of the lattice is kept constant during the assembly and expansion of the lattice, therefore curvature generation occurs from the very beginning of clathrin assembly (Figure 1b).

Schematic illustrations of the constant area model (a) and the constant curvature model (b) for CME. Blue: plasma membrane, yellow: clathrin coat, black dashed line: curvature of the clathrin coat

In order to distinguish between the two models, the dynamics of clathrin assembly and the geometry of membrane shapes are needed. Fluorescence microscopy, including light sheet and MINFLUX, has revealed the assembly dynamics of the clathrin coat (Aguet et al., 2016; Ferguson et al., 2016), as well as other proteins that participate in endocytosis (Sirotkin et al., 2010; Taylor et al., 2011; Kukulski et al., 2016; Kaksonen et al., 2006; Balzarotti et al., 2017; Gwosch et al., 2020; Schmidt et al., 2021), while electron tomography has been able to resolve membrane shapes during endocytosis (Avinoam et al., 2015; Kukulski et al., 2012). However, neither of the methods can capture both spatial and temporal information at the same time. Under conventional fluorescence microscopy, the clathrin-coated pits appear as diffraction-limited spots due to their small size, which is typically ∼ 30 − 150nm in mammals (McMahon and Boucrot, 2011) and yeast (Kukulski et al., 2012), and shape information of the membrane is completely lost (Cocucci et al., 2012; Loerke et al., 2009; Picco et al., 2015; Sirotkin et al., 2010; Kural and Kirchhausen, 2012). On the other hand, super-resolution fluorescence microscopy has been able to reveal the protein organization at the endocytic pit (Mund et al., 2018; Cocucci et al., 2012; Arasada et al., 2018) and to reconstruct the shape of the clathrin coat (Sochacki et al., 2017; Scott et al., 2018; Mund et al., 2023) from averaging over ensembles of endocytic sites. Correlative light and electron microscopy (CLEM) method has exploited the fluorescence of fiducial markers to locate endocytic sites while resolving membrane shapes using electron tomography. However, both super-resolution and CLEM requires sample fixation, therefore, one can identify multiple endocytic sites at the same time and perform the average, yet unable to trace a single endocytic site over time. The temporal information is nevertheless lost.

As a result of the incomplete information obtained by existing experimental methods, both hypotheses have experimental support. Experiments that apply electron microscopy to resolve the membrane shapes of endocytic pits favor the constant area model (Avinoam et al., 2015; Bucher et al., 2018; Sochacki and Taraska, 2019; Sochacki et al., 2021). However, super-resolution imaging combined with analysis of the fluorescence intensity of the clathrin coat is inclined towards the constant curvature model (Willy et al., 2021). In addition, it was argued that the energetic cost of bond reorganization in a regular hexagonal lattice in the constant area model may be too large to be fulfilled (Frey et al., 2020; Frey and Schwarz, 2020; Kirchhausen et al., 2014).

Extensive theoretical efforts have been dedicated to model membrane morphology during endocytosis (Fu and Johnson, 2023), of which molecular dynamics simulations (Varga et al., 2020) and continuum mechanics (Hassinger et al., 2017; Ma and Berro, 2021; Walani et al., 2015; Agrawal and Steigmann, 2008; Rangamani et al., 2013) are two common approaches. Hybrid models were also broadly applied to gain higher resolution than continuum mechanics and lower computing expense than molecular dynamics (Fu et al., 2019, 2021). However, most theoretical investigations have focused on how mechanical properties, such as membrane tension and bending rigidity of the clathrin coat, influence the membrane morphology. The process of curvature generation is either neglected or taken for granted. Only few of them have addressed the difference between the constant area model and the constant curvature model.

In fact, the constant curvature and constant area models are only two extreme models for clathrin assembly during endocytosis and any change in area or curvature are possible at any time point during endocytosis. In this paper, we extend the classic Helfrich theory for membrane deformation to incorporate energy terms associated with clathrin assembly and curvature generation, and compare geometric features calculated by theory with that extracted from experimental data. The negative gradient of the total free energy defines a pathway that neither fits the constant area model nor the constant curvature model. We find that a pathway that is close to the constant area model fits electron tomograms of the endocytic pits the best. Our study also offers experimental suggestions to distinguish between the two main hypotheses.

Models and methods

We model the membrane patch of the CCP (clathrin-coated pit) as a surface which is rotationally symmetric with respect to the z-axis. The shape of the membrane is parameterized with the meridional curve {r(s), z(s)}, where s denotes the arc length along the curve. The bending energy of the membrane (together with the clathrin coat) assumes the Helfrich model (Helfrich, 1973) where k denotes the bending rigidity of the CCP, C1 and C2 denote the principal curvatures of the surface, C0 denotes the intrinsic curvature of the membrane induced by the clathrin coat. To model a finite area of the clathrin coat, we assume the intrinsic curvature C0 spatially varies as where a denotes positions on the membrane. Here we choose a to be the surface area calculated from the tip of the membrane, and C0 equals c0 for area a < a0, and rapidly drops to zero when a > a0. The parameter α controls the sharpness of the drop. In the constant area model, we vary the intrinsic curvature c0 but keep the coating area a0 constant, while in the constant curvature model, we vary the coating area a0 but keep the intrinsic curvature c0 constant. As a result of clathrin coat, the bending rigidity k also varies as where kcoat and kbare denotes the bending rigidity of the clathrin-coated membrane and bare membrane, respectively. Besides the bending energy, the membrane tension contributes to the free energy in the form of where σe denotes the membrane tension at the base and A denotes the surface area of the membrane patch within a fixed radius of Rb. The membrane tension σe and bending rigidity k defines a characteristic length (Derényi et al., 2002). The total free energy Etot = Eb + Et is a functional of the membrane shape. Given a coating area a0 and an intrinsic curvature c0, we numerically solve the variational equations of the energy functional to obtain membrane shapes that minimizes Etot. More detailed descriptions of the model can be found in Appendix 1.

Results

Difference between the constant area model and the constant curvature model in terms of membrane morphology

Vesiculation requires assembly of a clathrin coat on the membrane, as well as curvature generation from the clathrin coat. A vesiculation process defines a pathway in the phase space (a0, c0) of the clathrin coat area a0 and the intrinsic curvature c0 of the coat. The constant area model and the constant curvature model are pathways that are made of a vertical line and a horizontal line. Besides these two extreme cases, there is a continuum spectrum of pathways with simultaneously increasing coating area a0 and intrinsic curvature c0 that could lead to vesiculation. Along the pathway, the membrane evolves from a flat shape to a dimple shape, and finally to an Ω-shape, as shown in Figure 2. Hereafter we use the maximal tangential angle Ψmax of the membrane as an indicator of the progression of vesiculation - when the membrane is flat, Ψmax = 0°, and when the membrane becomes spherical, Ψmax = 180°. In our simulation, the neck becomes extremely narrow before Ψmax = 180°. Therefore, we consider vesiculation occurs when Ψmax reaches 150° (Figure 2a).

Evolution of membrane morphology and phase diagram of vesiculation in the coating area vs. intrinsic curvature (a0, c0) parameter space. (a) Membrane shapes at different stages of invagination and definition of some variables used in this paper. We define the distance from the axisymmetric axis to the edge of the coating area as Rcoat, the radius of the tangential curvature circle at the tip of the shape as Rt, and the distance from axisymmetric axis to the boundary as Rb. The maximum tangential angle of the cross section contour is Ψmax. (b,c) Tip radius Rt vs. maximal angle Ψmax for the constant area model in (b) and for the constant curvature model in (c). Dotted lines in (b) denote the analytical solutions. Insets show the ratio of the tip radii Rt at Ψmax = 150° and Ψmax = 30°. The inset dark dots denotes the numerical results and the red line is the analytical solution (See Appendix 5). (d,e) Coat radius Rcoat vs. maximal angle Ψmax for the constant area model in (d) and for the constant curvature model in (e). Dotted lines in (d) denote the analytical solutions. Insets show the ratio of the coat radius Rcoat at Ψmax = 150° and Ψmax = 90°. The inset dark dots denotes the numerical results and the red lines are the analytical ones (See Appendix 5). (f) Vesiculation diagram in the phase space of (a0, c0). Each horizontal line represents a path of the constant curvature model and each vertical line represents a path of the constant area model. Each path terminates when Ψmax = 150°. The solid grey lines represent contours of Ψmax = 30°, 60°, 90°, 120°, 150°, respectively. The solid black line is the analytical results for the vesiculation line (See Appendix 4). The dashed black line is a random-picked straight line connecting the origin and the vesiculation boundary. The intersections of the dashed black line and the gray lines are plotted in orange dots and they are the coordinates of (a0, c0) where the five shapes in (a) are located. Shapes are arranged in an increasing order of Ψmax in (a). (b-e) Parameters with a bar over them (left vertical axes) are normalized to be dimensionless, and the dimensional parameters (right vertical axes) are calculated by one of the typical fitting values L0 = 40nm (Figure 5)

First, we analyze the difference between the two models in terms of membrane morphology evolution along their pathways. We fit a circle around the membrane tip and use the radius Rt of the circle to characterize the curvature of the membrane at the tip. When Rt is plotted against the maximal angle Ψmax, both models show that Rt decreases with increasing Ψmax (Figure 2b and c). We stress that even though the intrinsic curvature c0 is fixed in the constant curvature model, it does not imply the tip radius along the vesiculation pathway is a constant. When the coating area is small, the geometric curvature at the membrane tip remains small and differs from the intrinsic curvature. Tip radius in the constant curvature model decays more steeply with Ψmax than in the constant area model. This difference becomes obvious when one plots the ratio of the tip radius at Ψmax = 150° and Ψmax = 30° (Figure 2b and c insets). For the constant area model, the ratio approximately equals to a constant 0.268 regardless of the coating area a0, which agrees well with the analytical result (See Appendix 5), while for the constant curvature model, the ratio stays above and approaches 1 at a certain c0 value.

Another difference between the two models is the evolution of the projected area of the clathrin coat on the substrate. We use the maximal radius Rcoat of the membrane within the clathrin-coated area as the indicator of the projected area (Figure 2a). In the constant area model, Rcoat decreases with increasing Ψmax, while in the constant curvature model, Rcoat increases with Ψmax and reaches a plateau (Figure 2d and e). The ratio Rcoat (150°)/Rcoat (90°) is about 0.732 in the constant area model and around 1 in the constant curvature model (Figure 2d and e, insets). The analytical curves of Rt and Rcoat against Ψmax also fit perfectly with numerical solutions in the constant area model (Figure 2b and d, compare dotted and solid curves). Our calculations therefore demonstrate that the two models exhibit clear differences in the evolution of Rt and Rcoat at the beginning of endocytosis (i.e. when Ψmax is small) which can be determined from shapes of endocytic pits obtained experimentally.

To demonstrate how the coating area a0 and the intrinsic curvature c0 of the clathrin coat influence the membrane morphology, for each pair of (a0, c0), we calculate the corresponding membrane shapes and plot the contour lines for Ψmax which indicate the stage of endocytosis (Figure 2f). The contour line with Ψmax = 150° represents the critical line where vesiculation occurs. The line can be well fitted by the analytical expression (Figure 2f, thick black line, see Appendix 4). It implies that a small intrinsic curvature of the clathrin coat is able to induce vesiculation of the membrane with a large clathrin coat. We find that the distances between contour lines for higher values of Ψmax is smaller than those for lower values of Ψmax. It means that at the late stage of endocytosis, a small change in a0 and c0 could result in a more dramatic change in the membrane shape than that at the early stage. This trend is demonstrated clearly from the orange dots in Figure 2f, which correspond to shapes in Figure 2a.

Note that in order to produce the phase diagram (Figure 2f), it requires that the bending rigidity of the clathrin coated area kcoat is significantly larger than kbare in the uncoated area. If kcoat is comparable with kbare, there exists a region in the phase diagram in which a single (a0, c0) corresponds to three possible membrane shapes (See Appendix 2—Figure 1a-d). Physically, it implies a discontinuous transition in the membrane shape along a path that passes through this region, and a gap in the maximal angle Ψmax would appear. Because in experiments, a wide spectrum of Ψmax are observed and no gap in Ψmax is found (Avinoam et al., 2015), we keep kcoat much greater than kbare for the rest of the paper. In this regime, the membrane shapes evolve continuously along any pathway that connects the origin (0, 0) with a point on the critical vesiculation curve.

Vesiculation needs free energy sources to drive clathrin assembly and curvature generation

In the previous section, we take curvature generation in the constant area model and clathrin assembly in the constant curvature model for granted, so that the coating area a0 and the intrinsic curvature c0 are imposed, such that the physical forces behind clathrin assembly and curvature gen-eration are ignored. However, the bending energy Eb and the tension energy Et typically increase with a0 and c0 along a vesiculation pathway. Therefore, vesiculation will be energetically unfavorable if no additional free energy sources are provided. In this section, we extend the model to include free energy terms for the assembly of the clathrin coat and its reorganization for curvature generation.

To describe the assembly of the clathrin coat in the constant curvature model, we introduce where μ denotes the effective surface binding energy density of clathrin molecules with the membrane. This term reduces the free energy with increasing a0, therefore, driving the assembly of clathrin. We can identify three types of free energy curves for different assembly strength μ: (1) When μ is small, the total free energy Etot = Eb + Et + Ea as a function of the coating area a0 has two local minima, with the lowest one at a small a0 and the other one at the maximum a0 where vesiculation occurs (Figure 3a, red curve). The minima are separated by an energy barrier that is significantly higher than the thermal energy kBT and the clathrin coat would assemble to a small area and halt. (2) With increasing μ, the lowest free energy minimum is shifted to the vesicula-tion point, but the energy barrier still exists and the clathrin coat remains small (Figure 3a, orange curve). (3) Vesiculation could happen for large enough μ such that the energy barrier vanishes and the free energy Etot monotonically decreases with a0 (Figure 3a, green). Based on the above analysis of the energy landscape we construct the phase diagram of the constant curvature model with clathrin assembly in the phase space of (c0, μ) and classify the points into four types. Besides the three types mentioned above, when the intrinsic curvature c0 is small, increasing a0 to its maximum value (105nm2) cannot produce vesiculation. (Figure 3b, gray region). The critical assembly energy density μ at which the energy barrier vanishes is found to increase with the intrinsic curvature c0 (Figure 3b, interface between the green region and the orange region), which implies that a larger assembly strength of clathrin coat μ is needed to complete vesiculation if the clathrin coat has a higher intrinsic curvature c0. When comparing the contour lines of the energy barrier ΔEtot = 1kBT and ΔEtot = 10kBT (Figure 3b, dotted curve and dash-dotted curve), the gap between them increases with c0, which means that the energy efficiency is reduced with c0 in the sense that, for larger c0, a larger increase in μ is needed to reduce the same amount of free energy.

Free energy evolution in the constant curvature and constant area models when accounting for one of either the polymerization energy term Ea = −μa0 or the curvature generation energy term Ec = −va0c0 (type 1) and . (a,b) Free energy landscape of the modified constant curvature model where with polymerization energy Ea = −μa0 in (a) and the corresponding phase diagram in the phase space of (c0, μ) in (b). The rightmost endpoint of each curve is the vesiculation point where Ψmax = 150°. The red line in (a) and red dots in (b) correspond to pathways with minimum free energy Etot appearing at a point other than the vesiculation point on the Etot − ā0 curve. The orange line in (a) and the orange dots in (b) correpsond to pathways where the vesiculation point is the minimum free energy point, but an energy barrier still exists. The green line in (a) and green dots in (b) correspond to vesiculation pathways without an energy barrier. The energy barrier ΔEtot is defined as the energy difference between the maximum point and the first minimum point before the maximum. ΔEtot of the orange curve is shown in (a) as a typical example. The gray dots in the left-hand side of (b) correspond to pathways that numerically fail to reach the vesiculation point when ā0 reaches its upper limit 10. (c,d) Free energy landscape of the modified constant area model where ā0 = 2 with curvature generation energy Ec = −va0c0 in (c) as a function of the intrinsic curvature c0 and the corresponding phase diagram in the phase space of (a0, v) in (d). The gray dots in the left-hand side of (d) correspond to pathways that numerically fail to reach the vesiculation point when reaches its upper limit 5. (e,f) Free energy landscape of the modified constant area model where ā0 = 2 with in (e) and the corresponding phase diagram in the phase space of (a0, v) in (f). The gray dots in the left-hand side of (f) correspond to pathways that numerically fail to reach the vesiculation point when reaches its upper limit 5. (a-f) The parameters with a bar over them are normalized to be dimensionless, and the dimensional parameters are calculated using one of the typical fitting values L0 = 40nm (Figure 5). (b,d,f) The black line separates the region wiht gray dots (which did not numerically reach vesicultion) from the other regions. The dotted, dash-dotted and solid gray lines respectively represent the paths where ΔEtot = 1kBT, ΔEtot = 10kBT, ΔEtot = 100kBT.

We next consider curvature generation in the constant area model. As the molecular mechanisms of curvature generation of the clathrin coat remains debated, we introduce a phenomenological model in which the free energy has the general form, where v denotes the strength of curvature generation, m and n are two positive numbers that are associated with the molecular mechanisms of curvature generation. The free energy Ec in Equation 6 decreases with increasing c0, therefore, driving curvature generation. We set m = 1 such that Ec is proportional to the coating area. Note that m cannot be zero, otherwise, Ec only depends on the intrinsic curvature c0 and can be nonzero even when the coating area is zero. As for the power n of the intrinsic curvature c0, we set n = 1 or 2 (called Model(1,1) and Model(1,2), respectively). Physically, Model(1,2) implies cooperativity in the curvature generation such that the reduction of free energy per increase of unit curvature is proportional to the current curvature, i.e., ΔEc ∝ −c0Δc0, while in Model(1,1), the reduction of free energy per increase of unit curvature is independent of current curvature. For Model(1,1), when v is small, the total free energy Etot = Eb + Et + Ec as a function of the intrinsic curvature c0 has two minima, the lowest one at a small positive c0 and the other one at the maximum c0 where vesiculation occurs (Figure 3c, red curve). Further curvature generation is strictly limited by the high energy barrier (sometimes more than 100kBT) between the two minima. With increasing v, the lowest minimum shifts to the vesiculation point, but the energy barrier still prevents curvature generation (Figure 3c, orange line). For a large enough v, the free energy monotonically decreases with c0 and the curvature generation proceeds until vesiculation occurs (Figure 3c, green curve). When the coating area a0 is very small, vesiculation fails to occur even when the intrinsic curvature is increased to its maximum value (0.125nm−1) (Figure 3c, gray region). In the phase space of (a0, v), the critical value of v where the energy barrier vanishes increases with the coating area a0 (Figure 3d, interface between the orange region and the green region), which implies that a larger clathrin coat needs a stronger strength of curvature generation to complete vesiculation.

Model(1,2) has similar free energy landscape as Model(1,1) (Compare Figure 3c and e, d and f). However, in Model(1,2), for very small v, the lowest free energy minimum is strictly pinned at c0 = 0, which implies no spontaneous curvature generation. In contrast, the minimum is at a small positive c0 in Model(1,1), which indicates slight curvature generation.

Determination of the vesiculation pathway from the energy landscape

In this section, we combine the assembly energy Equation 5 and the curvature generation energy Equation 6 together and calculate the total free energy Etot (a0, c0) = Eb +Et +Ea +Ec as a function of both the coating area a0 and the intrinsic curvature c0. A pathway from the origin can be constructed by the descent along the negative gradient of the free energy landscape −∇Etot. In Figure 4a we show the free energy landscape for a fixed assembly strength and varied reorganization strength for model(1,2). When v is small, the energy contour lines near the origin are kinked, which represents an energy barrier that prevents the path from going up, i.e. from generating curvature. The path extends horizontally and terminates on the a0 − axis (Figure 4a, first column, red curve). With increasing v, the kinked contour lines shift towards larger a0 and the path can be lifted up to in the middle and drops to the a0 − axis in the end (Figure 4a, second column, orange curve). Beyond a critical v, the energy barrier vanishes and the path bends up and terminates on the vesiculation curve (Figure 4a, third column, green curve). Further increasing μ leads to the path lifting up at a smaller coating area a0 (Figure 4a, fourth column, green curve). The path goes horizontally first and is later lifted up, which resembles the path of the constant area model. The three types of pathways are classified into three colored regions in the phase diagram of (μ, v), which represent complete vesiculation (green), partial vesiculation (orange) and no vesiculation (red), respectively (Figure 4c).

Vesiculation phase diagram when accounting for both the polymerization energy μ and the reorganization energy v. (a,b) Free energy landscape for Model(1,2) (i.e. with reorganization energy in (a) and Model(1,1) (i.e. with reorganization energy in (b). Thick black lines are the analytical solutions for the vesiculation boundary. Thin rainbow-colored lines visually represent the energy landscape (values of the color bar are for the dimensionless free energy scale). Thick colored-lines represent pathways that stream along the negative gradient of the free energy landscape in the phase space starting from the origin (i.e. no clathrin assembled and no curvatuve). Our model shows that only a subset of suitable (μ, v) values create pathways that lead to vesiculation, i.e. that reach the thick black line (thick green lines in the third and fourth panels). The orange and red curves are pathways that fail to reach vesiculation. The red curve does not produce any curvature, while the orange curve generates a small curvature but never leads to vesiculation. (c,d) Phase diagrams for Model(1,2) in (c) and Model(1,1) in (d) show the relationship between pathway types and the (μ, v) values. The colors of the dots correspond to the same types of pathways as represented by thick colored lines in (a,b). Our results show that larger μ or v values lead to an easier vesiculation. Parameters with a bar over them are normalized to be dimensionless, and the dimensional parameters are calculated using one of the typical fitting values L0 = 40nm (Figure 5).

The free energy landscapes of Model(1,1) dramatically differs from Model(1,2). When v is small, the energy gradient is strongly biased towards the horizontal direction, and the path extends horizontally with little or no curvature generation (Figure 4b, first column). For an intermediate v, the path first goes towards the top right direction until and then slowly bends down and extends towards large coating area along a valley formed in the energy landscape, which corresponds to membrane shapes with a small dimple (Figure 4b, second column). For large enough v, the path shoots nearly straightly towards the top right direction before it reaches the vesiculation line (Figure 4b, third column). Further increasing v makes the path more straight and terminates at a smaller coating area (Figure 4b, fourth column). The (μ, v) phase diagram shows the parameter regions that lead to complete, partial or no vesiculation for Model(1,1) (Figure 4d).

Comparison between different models with the experimental data

The constant area model and the constant curvature model represent two extreme pathways of membrane vesiculation. We have found constant-area-like pathways in Model(1,2) and straight-line-like pathways in Model(1,1). In order to understand which model is the most plausible, we compare membrane shapes predicted by the models with the membrane profiles obtained by electron microscopy in (Avinoam et al., 2015). The fitting error ϵ of a vesiculation path reflects the relative difference between the model-predicted geometric features along the path and the rolling median of the corresponding experimental data (Figure 5 and Appendix 3). The fitting geometric features include neck width, tip radius and invagination depth (Figure 5c). We draw the corresponding optimum energy paths that minimize the fitting error (Figure 5b), and compare the best model-predicted shapes with the experimental ones (Figure 5d).

Comparison between our theory and experimental data from mammalian cells. (a) Parameter fit of the best of the four models (constant curvature model, constant area model, Model(1,1), Model(1,2)) to obtain the minimum error ϵ. Fitting procedure of Model(1,1) and Model(1,2) consider the total free energy Etot = Eb + Et + Ea + Ec, while the fitting of the constant area model and the constant curvature model consider Etot = Eb + Et. The optimized parameters are ā0 ∈ [0, 10] for the constant area model, for the constant curvature model, and for model(m,n), and L0 ∈ [10nm, 100nm] within an interval of 10nm in the four models. We only assign fitting errora to the parameter sets that lead to vesiculation and only plot the error figure for the best L0. (b) Vesiculation pathways with minimum fitting error in the four models. In each model, we use the best L0 value from (a) to obtain the dimensional scale of the (a0, c0) phase space. (c) Comparison of model fits and experimental data for three geometric features: neck width, tip radius (Rt) and invagination depth. Neck width is calculated as the distance between the left and right parts of the shape for Ψmax = 90°, and the invagination depth is measured as the height from the base to the tip of the invagination. (d) Comparison between the model-predicted shapes and the experimental shapes. Experimental membrane shapes for mammalian cells are grouped according to their maximum angle as a proxy for the different stages of CME. The number of experimental shapes falling in a certain Ψmax range is defined as n. The black lines are the average experimental shapes after symmetrization. The model-predicted shapes are calculated by the midpoint value of each Ψmax interval. (c,d) The curves predicted by theory are shown with colored lines, and experimental data is shown with gray dots and black lines. Parameters with a bar over them are non-dimensionalized. The detailed procedure to treat the experimental data can be found in Appendix 3.

For Model(1,2) and Model(1,1), the fitting parameters include the polymerization strength and the reorganization strength which together determine the vesiculation pathway, as well as the characteristic length L0 which scales the size of the membrane. For Model(1,2), the best fits are obtained for and L0 = 30nm (Figure 5a). The resulting path moves horizontally at first and then bents up vertically (Figure 5b), which resembles the behavior of the constant area model. For Model(1,1), the best fitting parameters are and L0 = 40nm (Figure 5a). The fitting path is almost a straight line towards the vesiculation line (Figure 5b). The optimum fitting error of Model(1,2) (ϵ = 0.14) is slightly better than that of Model(1,1) (ϵ = 0.17).

We also perform the fitting procedure to the constant area model and find the optimum parameters are a0 = 1.69 × 104nm2 and L0 = 30nm. For the constant curvature model, the best fitting parameters are c0 = 0.043nm−1 and L0 = 50nm. The minimum fitting error of the constant curvature model (ϵ = 0.28) is exactly twice as large as that of the constant area model (ϵ = 0.14) (Figure 5a and b). So considering the fitting error and the pathway in (a0, c0) phase diagram, we raise the conclusion that the experimental vesiculation process probably favors constant-area-like pathways.

When comparing the model-predicted geometric features with the rolling median of the experimental data, we find that the four models fit almost equally well the experimental data for the neck width(Figure 5c left). Model(1,2) and the constant area model predict very similar results such that the curves almost overlap with each other (Figure 5c, red curve and orange curve). The predictions of these two models match the rolling median of the experimental data best. The constant curvature model strongly deviates from the rolling median of the experimental tip radius, particularly in the early stage of vesiculation when Ψmax < 90°.

To compare the axisymmetric membrane shapes predicted by the models with the non-axisymmetric membrane profiles obtained with electron microscopy, we symmetrize the experimental data with a procedure (Appendix 3—Figure 1). Then, we average the symmetrized profile within an interval of Ψmax ∈ [Ψ0 −10°, Ψ0 +10°] and overlay the averaged profile with model-predicted shapes for Ψmax = Ψ0 (Figure 5d). At the early stage when the membrane exhibits a dimple shape (0° < Ψmax < 60°), the membrane morphology predicted by the constant curvature model is distinguishable from the other three models, particularly when looking at the tip radius. At the late stage when the membrane exhibits an Ω-shape, i.e., Ψmax > 90°, the difference in shape between models is mainly manifested in the invagination depth. The constant curvature model and Model(1,1) mainly predict a deeper invagination depth than the symmetrized experimental profile, while the constant area model and Model(1,2) usually give much better fitting.

Discussion

Three types of clathrin coats

In this paper, we have constructed a physical model to describe how curvature generation and clathrin assembly are interrelated during the vesiculation process in CME. Previous experiments have reported three groups of clathrin coated pits, which are plaques, abortive pits and pits that lead to vesiculation, according to their structural and dynamic properties (Maupin and Pollard, 1983; Ehrlich et al., 2004; Loerke et al., 2009; Saffarian and Kirchhausen, 2009; Kirch-hausen, 2009; Lampe et al., 2016).

In Figure 4, we show that depending on the clathrin assembly strength μ and its reorganization strength v, the pathway might end up with three possible final shapes: (i) a flat membrane with no curvature generation, (ii) a nearly flat membrane with small curvature generation, (iii) a spherical cap that leads to vesiculation. They essentially correspond to the three types of clathrincoated pits found in experiments. Based on the phase diagram of Model(1,2) (Figure 4c), the difference between the three groups comes from the difference in the assembly and reorganization strengths of clathrin molecules. Furthermore, Model(1, 2) predicts that at the boundary between the type (iii) region and the type (ii) region, the reorganization strength v increases with the assembly strength. This result has important implications to explain why large plaques are commonly observed in experiments. The large area of the plaques are due to the strong assembly strength μ. However, for these plaques to go to vesiculation, a strong reorganization energy v is also needed. Therefore, the combination of a strong μ and weak v leads to the formation of large plaques. Model(1,2) predicts that a plaque or an abortive pit can be transformed into a vesicle by either increasing the reorganization strength or reducing the assembly strength (Figure 6). The former ends up with a large vesicle and the latter ends up with a small vesicle. This can be used as a test of our theory with experiments to modify the binding affinity of clathrin molecules with adaptor proteins on the membrane. Weakening the affinity might increase the portion of vesicles and reduce the portion of plaques, though the vesicles would become smaller.

Tip radius of vesiculaion shapes (Rves) in Model(1,2). The colored region shows the (μ, v) sets that lead to vesiculation, and brighter colors correspond to larger Rves. Decreasing assembly strength μ or increasing reorganization strength v might lead to vesiculation of different vesicle sizes. An example from (μ, v) = (13.3 × 10−3kBT · nm−2, 8.8kBT) to the vesiculation region is marked by arrows, red dots and corresponding vesicle shapes. The characteristic length L0 = 30nm is used in the calculation.

Cooperativity in the curvature generation process

In Figure 5, we show the best fitting results for all the four models and find that Model(1,2) produces better fits than Model(1,1), which suggests the existence of cooperativity in the curvature generation process. In particular, if curvature generation is driven by breaking bonds in the hexagonal lattice, cooperativity implies that the number of newly broken bonds is proportional to the number of already broken bonds. Because of this cooperativity, at the early stage of endocytosis bonds are broken slowly and clathrin assembly dominates over curvature generation. At the late stage of endocytosis, an increasing number of bonds are broken and curvature generation could happen rapidly and dominate over clathrin assembly. Altogether, this cooperativity leads to a constant-area-like behavior. Similar effect have been reported in (Mund et al., 2023).

The difference in membrane morphology between the different models is most salient at the early stage of endocytosis

When we compare the model predictions, we find that the difference in membrane morphology between models is not as big as expected, which might explain why it has been difficult to distinguish between the constant area and the constant curvature models for so long. For instance, the neck width vs. Ψmax and the invagination depth vs. Ψmax are similar for all the four models (Figure 5c, left and right). The best fitting error of the four models calculated from the geometric features are relatively close, except for the constant curvature model, which gives the worst fit (Figure 5a). The models are mainly distinguishable from the tip radius vs. Ψmax plot at the early stage of endocytosis when the membrane is nearly flat (Figure 5c, middle), i.e. for shapes with small Ψmax. However, published experimental shape at early stages of endocytosis are sparse. Our result hints that in order to distinguish between the models, collecting membrane shapes at the early stage is necessary and the relation of tip radius vs. Ψmax is the key geometric feature to tell the models apart.

The projected area of clathrin coat in the plane of the plasma membrane could distinguish between the two models

In Figure 2d and e we have shown that the coat radius Rcoat, which represents the projected area of the clathrin coat in the plane of the plasma membrane, as a function of Ψmax exhibit opposite trends for the constant curvature model and the constant area model. The results suggest that in experiments the projected area for the constant area model would first increase and then decrease over time, finally reaching a plateau. However, for the constant curvature model, the projected area would increase over time and finally reach a plateau without a decreasing phase. This result suggests another method to distinguish between the two models via the projected area measurement. The idea has been used in a study where the clathrin-coated pit was imaged with platinum replica and cryoelectron microscopy and tomography (Sochacki et al., 2021). The results support a constant-area-like model, consistent with the prediction of our Model(1,2), in which the dome structures have a slightly larger coating area than flat structures. On the other hand, another study has used the super-resolved live cell fluorescence imaging with TIRF to measure the growth of the clathrin coat over time (Willy et al., 2021). The authors found a smooth drop in the projected area of clathrin coat over time. However, based on a computer simulation of clathrin assembly, they concluded that the smooth drop of the projected area was the result of a constant-curvature-like model because a constant-area-like model would produce a sharp drop. We attribute the difference between their model and our model to the fact that they model the clathrin coat as a discrete lattice while we use a continuum mechanics method. More importantly, in their model, the moment at which curvature generation occurs was arbitrarily imposed at 80% of clathrin triskelions. If the transition were chosen to occur with fewer triskelions, e.g. 40%, the sharp drop in the project area might not happen in the constant-area-like model. Furthermore, the authors used the number of triskelions to monitor the progress of endocytosis which terminates when the triskelions reach the maximum number. This choice might bias towards the constant-curvature-like model because the vesiculation may not happen at all when the clathrin numbers reaches its maximum.

The bending rigidity of the coated area should be much larger than the uncoated area

Comparison of our model to experimental data demonstrates that the relative bending rigidities of the coat and the membrane are constrained. Indeed, if kcoat /kbare < 50, the model predicts an abrupt change (or a gap) in Ψmax at the end of vesiculation (See Appendix 2—Figure 1), i.e., a snap-through transition (Hassinger et al., 2017). If a gap in Ψmax existed, we would expect that the distribution of experimental shapes to be discontinuous, with no or very few data corresponding to a certain range of Ψ. However, in the experiments (Avinoam et al., 2015), the endocytic pits shapes are continuously distributed across the Ψmax spectrum, indicating the ergodicity of Ψmax during the endocytic process. Our calculation suggests that the clathrin coat is about 50 times stiffer than the membrane.

eLife Assessment

This valuable study proposes a theoretical model of clathrin coat formation based on membrane elasticity that seeks to determine whether this process occurs by increasing the area of a protein-coated patch with constant curvature, or by increasing the curvature of a protein-coated patch that forms in an initially flat conformation (so called constant curvature or constant area models). Identifying energetically favorable pathways and comparing the obtained shapes with experiments provides solid support to the constant-area pathway. This work will be of interest for biologists and biophysicists interested in membrane remodelling and endocytosis. It provides an innovative approach to tackle the question of constant curvature vs. constant area coat protein formation, although some of the model's assumption are only partially supported by experimental evidence.

Reviewer #1 (Public review):

Summary:

The authors develop a set of biophysical models to investigate whether a constant area hypothesis or a constant curvature hypothesis explains the mechanics of membrane vesiculation during clathrin-mediated endocytosis.

Strengths:

The models that the authors choose are fairly well-described in the field and the manuscript is well-written.

Weaknesses:

One thing that is unclear is what is new with this work. If the main finding is that the differences are in the early stages of endocytosis, then one wonders if that should be tested experimentally. Also, the role of clathrin assembly and adhesion are treated as mechanical equilibrium but perhaps the process should not be described as equilibria but rather a time-dependent process. Ultimately, there are so many models that address this question that without direct experimental comparison, it's hard to place value on the model prediction.

While an attempt is made to do so with prior published EM images, there is excessive uncertainty in both the data itself as is usually the case but also in the methods that are used to symmetrize the data. This reviewer wonders about any goodness of fit when such uncertainty is taken into account.

Reviewer #2 (Public review):

Summary:

In this manuscript, the authors employ theoretical analysis of an elastic membrane model to explore membrane vesiculation pathways in clathrin-mediated endocytosis. A complete understanding of clathrin-mediated endocytosis requires detailed insight into the process of membrane remodeling, as the underlying mechanisms of membrane shape transformation remain controversial, particularly regarding membrane curvature generation. The authors compare constant area and constant membrane curvature as key scenarios by which clathrins induce membrane wrapping around the cargo to accomplish endocytosis. First, they characterize the geometrical aspects of the two scenarios and highlight their differences by imposing coating area and membrane spontaneous curvature. They then examine the energetics of the process to understand the driving mechanisms behind membrane shape transformations in each model. In the latter part, they introduce two energy terms: clathrin assembly or binding energy, and curvature generation energy, with two distinct approaches for the latter. Finally, they identify the energetically favorable pathway in the combined scenario and compare their results with experiments, showing that the constant-area pathway better fits the experimental data.

Strengths:

The manuscript is well-written, well-organized, and presents the details of the theoretical analysis with sufficient clarity.

The calculations are valid, and the elastic membrane model is an appropriate choice for addressing the differences between the constant curvature and constant area models.

The authors' approach of distinguishing two distinct free energy terms-clathrin assembly and curvature generation-and then combining them to identify the favorable pathway is both innovative and effective in addressing the problem.

Notably, their identification of the energetically favorable pathways, and how these pathways either lead to full endocytosis or fail to proceed due to insufficient energetic drives, is particularly insightful.

Weaknesses:

Membrane remodeling in cellular processes is typically studied in either a constant area or constant tension ensemble. While total membrane area is preserved in the constant area ensemble, membrane area varies in the constant tension ensemble. In this manuscript, the authors use the constant tension ensemble with a fixed membrane tension, σe. However, they also use a constant area scenario, where 'area' refers to the surface area of the clathrin-coated membrane segment. This distinction between the constant membrane area ensemble and the constant area of the coated membrane segment may cause confusion.

As mentioned earlier, the theoretical analysis is performed in the constant membrane tension ensemble at a fixed membrane tension. The total free energy E_tot of the system consists of membrane bending energy E_b and tensile energy E_t, which depends on membrane tension, σe. Although the authors mention the importance of both E_b and E_t, they do not present their individual contributions to the total energy changes. Comparing these contributions would enable readers to cross-check the results with existing literature, which primarily focuses on the role of membrane bending rigidity and membrane tension.

The authors introduce two different models, (1,1) and (1,2), for generating membrane curvature. Model 1 assumes a constant curvature growth, corresponding to linear curvature growth, while Model 2 relates curvature growth to its current value, resembling exponential curvature growth. Although both models make physical sense in general, I am concerned that Model 2 may lead to artificial membrane bending at high curvatures. Normally, for intermediate bending, ψ > 90, the bending process is energetically downhill and thus proceeds rapidly. the bending process is energetically downhill and thus proceeds rapidly. However, Model 2's assumption would accelerate curvature growth even further. This is reflected in the endocytic pathways represented by the green curves in the two rightmost panels of Fig. 4a, where the energy steeply increases at large ψ. I believe a more realistic version of Model 2 would require a saturation mechanism to limit curvature growth at high curvatures.

Author response:

The following is the authors’ response to the original reviews

Reviewer #1:

Summary

The authors develop a set of biophysical models to investigate whether a constant area hypothesis or a constant curvature hypothesis explains the mechanics of membrane vesiculation during clathrin-mediated endocytosis.

Strengths

The models that the authors choose are fairly well-described in the field and the manuscript is wellwritten.

Thank you for your positive comments on our work.

Weaknesses

One thing that is unclear is what is new with this work. If the main finding is that the differences are in the early stages of endocytosis, then one wonders if that should be tested experimentally. Also, the role of clathrin assembly and adhesion are treated as mechanical equilibrium but perhaps the process should not be described as equilibria but rather a time-dependent process. Ultimately, there are so many models that address this question that without direct experimental comparison, it's hard to place value on the model prediction.

Thank you for your insightful questions. We fully agree that distinguishing between the two models should ultimately be guided by experimental tests. This is precisely the motivation for including Fig. 5 in our manuscript, where we compare our theoretical predictions with experimental data. In the middle panel of Fig. 5, we observe that the predicted tip radius as a function of 𝜓𝑚𝑎𝑥 from the constant curvature model (magenta curve) deviates significantly from both the experimental data points and the rolling median, highlighting the inconsistency of this model with the data.

Regarding our treatment of clathrin assembly and membrane adhesion as mechanical equilibrium processes, our reasoning is based on a timescale separation argument. Clathrin assembly typically occurs over approximately 1 minute. In contrast, the characteristic relaxation time for a lipid membrane to reach mechanical equilibrium is given by , where 𝜇∼5 × 10-9 𝑁𝑠𝑚-1 is the membrane viscosity, 𝑅0 =50𝑛𝑚 is the vesicle size, 𝜅=20 𝑘𝐵𝑇 is the bending rigidity. This yields a relaxation time of 𝜏≈1.5 × 10−4𝑠, which is several orders of magnitude shorter than the timescale of clathrin assembly. Therefore, it is reasonable to treat the membrane shape as being in mechanical equilibrium throughout the assembly process.

We believe the value of our model lies in the following key novelties:

(1) Model novelty: We introduce an energy term associated with curvature generation, a contribution that is typically neglected in previous models.

(2) Methodological novelty: We perform a quantitative comparison between theoretical predictions and experimental data, whereas most earlier studies rely on qualitative comparisons.

(3) Results novelty: Our quantitative analysis enables us to unambiguously exclude the constant curvature hypothesis based on time-independent electron microscopy data.

In the revised manuscript (line 141), we have added a statement about why we treat the clathrin assembly as in mechanical equilibrium.

While an attempt is made to do so with prior published EM images, there is excessive uncertainty in both the data itself as is usually the case but also in the methods that are used to symmetrize the data. This reviewer wonders about any goodness of fit when such uncertainty is taken into account.

Author response: We thank the reviewer for raising this important point. We agree that there is uncertainty in the experimental data. Our decision to symmetrize the data is based on the following considerations:

(1) The experimental data provide a one-dimensional membrane profile corresponding to a cross-sectional view. To reconstruct the full two-dimensional membrane surface, we must assume rotational symmetry.

(2)In addition to symmetrization, we also average membrane profiles within a certain range of 𝜓𝑚𝑎𝑥 values (see Fig. 5d). This averaging helps reduce the uncertainty (due to biological and experimental variability) inherent to individual measurements.

(3)To further address the noise in the experimental data, we compare our theoretical predictions not only with individual data points but also with a rolling median, which provides a smoothed representation of the experimental trends.

These steps are taken to ensure a more robust and meaningful comparison between theory and experiments.

In the revised manuscript (line 338), we have explained why we have to symmetrize the data:

“To facilitate comparison between the axisymmetric membrane shapes predicted by the model and the non-axisymmetric profiles obtained from electron microscopy, we apply a symmetrization procedure to the experimental data, which consist of one-dimensional membrane profiles extracted from cross-sectional views, as detailed in Appendix 3 (see also Appendix 3--Fig. 1).”

Reviewer #2:

Summary

In this manuscript, the authors employ theoretical analysis of an elastic membrane model to explore membrane vesiculation pathways in clathrin-mediated endocytosis. A complete understanding of clathrin-mediated endocytosis requires detailed insight into the process of membrane remodeling, as the underlying mechanisms of membrane shape transformation remain controversial, particularly regarding membrane curvature generation. The authors compare constant area and constant membrane curvature as key scenarios by which clathrins induce membrane wrapping around the cargo to accomplish endocytosis. First, they characterize the geometrical aspects of the two scenarios and highlight their differences by imposing coating area and membrane spontaneous curvature. They then examine the energetics of the process to understand the driving mechanisms behind membrane shape transformations in each model. In the latter part, they introduce two energy terms: clathrin assembly or binding energy, and curvature generation energy, with two distinct approaches for the latter. Finally, they identify the energetically favorable pathway in the combined scenario and compare their results with experiments, showing that the constant-area pathway better fits the experimental data.

Thank you for your clear and comprehensive summary of our work.

Strengths

The manuscript is well-written, well-organized, and presents the details of the theoretical analysis with sufficient clarity. The calculations are valid, and the elastic membrane model is an appropriate choice for addressing the differences between the constant curvature and constant area models.

The authors' approach of distinguishing two distinct free energy terms-clathrin assembly and curvature generation-and then combining them to identify the favorable pathway is both innovative and effective in addressing the problem.

Notably, their identification of the energetically favorable pathways, and how these pathways either lead to full endocytosis or fail to proceed due to insufficient energetic drives, is particularly insightful.

Thank you for your positive remarks regarding the innovative aspects of our work.

Weaknesses and Recommendations

Weakness: Membrane remodeling in cellular processes is typically studied in either a constant area or constant tension ensemble. While total membrane area is preserved in the constant area ensemble, membrane area varies in the constant tension ensemble. In this manuscript, the authors use the constant tension ensemble with a fixed membrane tension, σe. However, they also use a constant area scenario, where 'area' refers to the surface area of the clathrin-coated membrane segment. This distinction between the constant membrane area ensemble and the constant area of the coated membrane segment may cause confusion.

Recommendation: I suggest the authors clarify this by clearly distinguishing between the two concepts by discussing the constant tension ensemble employed in their theoretical analysis.

Thank you for raising this question.

In the revised manuscript (line 136), we have added a sentence, emphasizing the implication of the term “constant area model”:

“We emphasize that the constant area model refers to the assumption that the clathrin-coated area 𝑎0 remains fixed. Meanwhile, the membrane tension 𝜎𝑒 at the base is held constant, allowing the total membrane area 𝐴𝐴 to vary in response to deformations induced by the clathrin coat.”

Weakness: As mentioned earlier, the theoretical analysis is performed in the constant membrane tension ensemble at a fixed membrane tension. The total free energy E_tot of the system consists of membrane bending energy E_b and tensile energy E_t, which depends on membrane tension, σe. Although the authors mention the importance of both E_b and E_t, they do not present their individual contributions to the total energy changes. Comparing these contributions would enable readers to cross-check the results with existing literature, which primarily focuses on the role of membrane bending rigidity and membrane tension.

Recommendation: While a detailed discussion of how membrane tension affects their results may fall outside the scope of this manuscript, I suggest the authors at least discuss the total membrane area variation and the contribution of tensile energy E_t for the singular value of membrane tension used in their analysis.

Thank you for the insightful suggestion. In the revised manuscript (line 916), we have added Appendix 6 and a supplementary figure to compare the bending energy 𝐸𝑏 and the tension energy 𝐸𝑡. Our analysis shows that both energy components exhibit an energy barrier between the flat and vesiculated membrane states, with the tension energy contributing more significantly than the bending energy.

In the revised manuscript (line 151), we have also added one paragraph explaining why we set the dimensionless tension . This choice is motivated by our use of the characteristic length as the length scale, and as the energy scale. In this way, the dimensionless tension energy is written as

Where is the dimensionless area.

Weakness: The authors introduce two different models, (1,1) and (1,2), for generating membrane curvature. Model 1 assumes a constant curvature growth, corresponding to linear curvature growth, while Model 2 relates curvature growth to its current value, resembling exponential curvature growth. Although both models make physical sense in general, I am concerned that Model 2 may lead to artificial membrane bending at high curvatures. Normally, for intermediate bending, ψ > 90, the bending process is energetically downhill and thus proceeds rapidly. The bending process is energetically downhill and thus proceeds rapidly. However, Model 2's assumption would accelerate curvature growth even further. This is reflected in the endocytic pathways represented by the green curves in the two rightmost panels of Fig. 4a, where the energy steeply increases at large ψ. I believe a more realistic version of Model 2 would require a saturation mechanism to limit curvature growth at high curvatures.

Recommendation 1: I suggest the authors discuss this point and highlight the pros and cons of Model 2. Specifically, addressing the potential issue of artificial membrane bending at high curvatures and considering the need for a saturation mechanism to limit excessive curvature growth. A discussion on how Model 2 compares to Model 1 in terms of physical relevance, especially in the context of high curvature scenarios, would provide valuable insights for the reader.

Thank you for raising the question of excessive curvature growth in our models and the constructive suggestion of introducing a saturation mechanism. In the revised manuscript (line 405), following your recommendation, we have added a subsection “Saturation effect at high membrane curvatures” in the discussion to clarify the excessive curvature issue and a possible way to introduce a saturation mechanism:

“Note that our model involves two distinct concepts of curvature growth. The first is the growth of imposed curvature — referred to here as intrinsic curvature and denoted by the parameter 𝑐0 — which is driven by the reorganization of bonds between clathrin molecules within the coat. The second is the growth of the actual membrane curvature, reflected by the increasing value of 𝜓𝑚𝑎𝑥.

The latter process is driven by the former.

Models (1,1) and (1,2) incorporate energy terms (Equation 6) that promote the increase of intrinsic curvature 𝑐0, which in turn drives the membrane to adopt a more curved shape (increasing 𝜓𝑚𝑎𝑥). In the absence of these energy contributions, the system faces an energy barrier separating a weakly curved membrane state (low 𝜓𝑚𝑎𝑥) from a highly curved state (high 𝜓𝑚𝑎𝑥). This barrier can be observed, for example, in the red curves of Figure 3(a–c) and in Appendix 6—Figure 1. As a result, membrane bending cannot proceed spontaneously and requires additional energy input from clathrin assembly.

The energy terms described in Equation 6 serve to eliminate this energy barrier by lowering the energy difference between the uphill and downhill regions of the energy landscape. However, these same terms also steepen the downhill slope, which may lead to overly aggressive curvature growth.

To mitigate this effect, one could introduce a saturation-like energy term of the form:

where 𝑐𝑠 represents a saturation curvature. Importantly, adding such a term would not alter the conclusions of our study, since the energy landscape already favors high membrane curvature (i.e., it is downward sloping) even without the additional energy terms. “

Recommendation 2: Referring to the previous point, the green curves in the two rightmost panels of Fig. 4a seem to reflect a comparison between slow and fast bending regimes. The initial slow vesiculation (with small curvature growth) in the left half of the green curves is followed by much more rapid curvature growth beyond a certain threshold. A similar behavior is observed in Model 1, as shown by the green curves in the two rightmost panels of Fig. 4b. I believe this transition between slow and fast bending warrants a brief discussion in the manuscript, as it could provide further insight into the dynamic nature of vesiculation.

Thank you for your constructive suggestion regarding the transition between slow and fast membrane bending. As you pointed out, in both Fig. 4a (model (1,2)) and Fig. 4b (model (1,1)), the green curves tend to extend vertically at the late stage. This suggests a significant increase in 𝑐0 on the free energy landscape. However, we remain cautious about directly interpreting this vertical trend as indicative of fast endocytic dynamics, since our model is purely energetic and does not explicitly incorporate kinetic details. Meanwhile, we agree with your observation that the steep decrease in free energy along the green curve could correspond to an acceleration in dynamics. To address this point, we have added a paragraph in the revised manuscript (in Subsection “Cooperativity in the curvature generation process”) discussing this potential transition and its consistency with experimental observations (line 395):

“Furthermore, although our model is purely energetic and does not explicitly incorporate dynamics, we observe in Figure 3(a) that along the green curve—representing the trajectory predicted by model (1,2)—the total free energy (𝐸𝑡𝑜𝑡) exhibits a much sharper decrease at the late stage (near the vesiculation line) compared to the early stage (near the origin). This suggests a transition from slow to fast dynamics during endocytosis. Such a transition is consistent with experimental observations, where significantly fewer number of images with large 𝜓𝑚𝑎𝑥 are captured compared to those with small 𝜓𝑚𝑎𝑥 (Mund et al., 2023).”

The geometrical properties of both the constant-area and constant-curvature scenarios, as well depicted in Fig. 1, are somewhat straightforward. I wonder what additional value is presented in Fig. 2. Specifically, the authors solve differential shape equations to show how Rt and Rcoat vary with the angle ψ, but this behavior seems predictable from the simple schematics in Fig. 1. Using a more complex model for an intuitively understandable process may introduce counter-intuitive results and unnecessary complications, as seen with the constant-curvature model where Rt varies (the tip radius is not constant, as noted in the text) despite being assumed constant. One could easily assume a constant-curvature model and plot Rt versus ψ. I wonder What is the added value of solving shape equations to measure geometrical properties, compared to a simpler schematic approach (without solving shape equations) similar to what they do in App. 5 for the ratio of the Rt at ψ=30 and 150.

Thank you for raising this important question. While simple and intuitive theoretical models are indeed convenient to use, their validity must be carefully assessed. The approximate model becomes inaccurate when the clathrin shell significantly deviates from its intrinsic shape, namely a spherical cap characterized by intrinsic curvature 𝑐0. As shown in the insets of Fig. 2b and 2c (red line and black points), our comparison between the simplified model and the full model demonstrates that the simple model provides a good approximation under the constant-area constraint. However, it performs poorly under the constant-curvature constraint, and the deviation between the full model and the simplified model becomes more pronounced as 𝑐0 increases.

In the revised manuscript, we have added a sentence emphasizing the discrepancy between the exact calculation with the idealized picture for the constant curvature model (line 181):

“For the constant-curvature model, the ratio remains close to 1 only at small values of 𝑐0, as expected from the schematic representation of the model in Figure 1. However, as 𝑐0 increases, the deviation from this idealized picture becomes increasingly pronounced.”

Recommendation: The clathrin-mediated endocytosis aims at wrapping cellular cargos such as viruses which are typically spherical objects which perfectly match the constant-curvature scenario. In this context, wrapping nanoparticles by vesicles resembles constant-curvature membrane bending in endocytosis. In particular analogous shape transitions and energy barriers have been reported (similar to Fig.3 of the manuscript) using similar theoretical frameworks by varying membrane particle binding energy acting against membrane bending:

DOI: 10.1021/la063522m

DOI: 10.1039/C5SM01793A

I think a short comparison to particle wrapping by vesicles is warranted.

Thank you for your constructive suggestion to compare our model with particle wrapping. In the revised manuscript (line 475), we have added a subsection “Comparison with particle wrapping” in the discussion:

“The purpose of the clathrin-mediated endocytosis studied in our work is the recycling of membrane and membrane-protein, and the cellular uptake of small molecules from the environment — molecules that are sufficiently small to bind to the membrane or be encapsulated within a vesicle. In contrast, the uptake of larger particles typically involves membrane wrapping driven by adhesion between the membrane and the particle, a process that has also been studied previously (Góźdź, 2007; Bahrami et al., 2016). In our model, membrane bending is driven by clathrin assembly, which induces curvature. In particle wrapping, by comparison, the driving force is the adhesion between the membrane and a rigid particle. In the absence of adhesion, wrapping increases both bending and tension energies, creating an energy barrier that separates the flat membrane state from the fully wrapped state. This barrier can hinder complete wrapping, resulting in partial or no engulfment of the particle. Only when the adhesion energy is sufficiently strong can the process proceed to full wrapping. In this context, adhesion plays a role analogous to curvature generation in our model, as both serve to overcome the energy barrier. If the particle is spherical, it imposes a constant-curvature pathway during wrapping. However, the role of clathrin molecules in this process remains unclear and will be the subject of future investigation.”

Minor points:

Line 20, abstract, "....a continuum spectrum ..." reads better.

Line 46 "...clathrin results in the formation of pentagons ...." seems Ito be grammatically correct.

Line 106, proper citation of the relevant literature is warranted here.

Line 111, the authors compare features (plural) between experiments and calculations. I would write "....compare geometric features calculated by theory with those ....".

Line 124, "Here, we choose a ..." (with comma after Here).

Line 134, "The membrane tension \sigma_e and bending rigidity \kappa define a ...."

Line 295, "....tip radius, and invagination ...." (with comma before and).

Line 337, "abortive tips, and ..." (with comma before and).

We thank you for your thorough review of our manuscript and have corrected all the issues raised.

Version 2

Abstract

During clathrin-mediated endocytosis, a patch of flat plasma membrane is internalized to form a vesicle. In mammalian cells, how the clathrin coat deforms the membrane into a vesicle remains unclear and two main hypotheses have been debated. The “constant area” hypothesis assumes that clathrin molecules initially form a flat lattice on the membrane and deform the membrane by changing its intrinsic curvature while keeping the coating area constant. The alternative “constant curvature” hypothesis assumes that the intrinsic curvature of the clathrin lattice remains constant during the formation of a vesicle while the surface area it covers increases. Previous experimental studies were unable to unambiguously determine which hypothesis is correct. In this paper, we show that these two hypotheses are only two extreme cases of a continuum spectrum if we account for the free energies associated with clathrin assembly and curvature generation. By tracing the negative gradient of the free energy, we define vesiculation pathways in the phase space of the coating area and the intrinsic curvature of clathrin coat. Our results show that, overall, the differences in measurable membrane morphology between the different models are not as big as expected, and the main differences are most salient at the early stage of endocytosis. Furthermore, the best fitting pathway to experimental data is not compatible with the constant-curvature model and resembles a constant-area-like pathway where the coating area initially expands with minor changes in the intrinsic curvature, later followed by a dramatic increase in the intrinsic curvature and minor change in the coating area. Our results also suggest that experimental measurement of the tip radius and the projected area of the clathrin coat will be the key to distinguish between models.

Introduction

Clathrin-mediated endocytosis (CME) is a fundamental cellular process to transport lipids, membrane proteins and extracellular cargo molecules into the cell (McMahon and Boucrot, 2011; Sorkin and Puthenveedu, 2013; Lu et al., 2016; Kaksonen and Roux, 2018; Lacy et al., 2018; Mettlen et al., 2018). In mammalian cells, a small patch of flat plasma membrane is shaped into a spherical vesicle when CME occurs (Avinoam et al., 2015). Clathrin molecules are essential for the membrane remodelling process. They are made of three subunits that form a triskelion, which further assemble into a cage-like structure in vitro (Musacchio et al., 1999; Shraiman, 1997). The minimum cages contain 16 polygons (Fotin et al., 2004) and the most commonly observed ones are semi-regular icosahedral cages (Cheng et al., 2007; Dannhauser and Ungewickell, 2012; Heuser, 1980). Two main hypotheses are under debate regarding how the clathrin coat scaffolds the flat membrane into a spherical vesicle in vivo (Avinoam et al., 2015; Chen and Schmid, 2020; Frey and Schwarz, 2020; Kaksonen and Roux, 2018; Scott et al., 2018). The constant area model asserts that the clathrin molecules initially polymerize into a flat lattice with a regularly arranged hexagonal structure, and later reorganization of the bonds between adjacent clathrins results in the formation of pentagons in the hexagonal lattice, which in turn leads to curvature generation of the clahtrin coat (Fotin et al., 2004; Heuser and Anderson, 1989) (Figure 1a). Adhesion of clathrin molecules with the substrate has been suggested to contribute a flattening force that prevents curvature generation. Release of the flattening force therefore could induce curvature of the clathrin coat with preloaded pentagons (Sochacki et al., 2021). The alternative constant curvature model asserts that the intrinsic curvature of the elements of the lattice is kept constant during the assembly and expansion of the lattice, therefore curvature generation occurs from the very beginning of clathrin assembly (Figure 1b).

Schematic illustrations of the constant area model (a) and the constant curvature model (b) for CME.

Blue: plasma membrane, yellow: clathrin coat, black dashed line: curvature of the clathrin coat.

In order to distinguish between the two models, the dynamics of clathrin assembly and the geometry of membrane shapes are needed. Fluorescence microscopy, including light sheet and MINFLUX, has revealed the assembly dynamics of the clathrin coat (Aguet et al., 2016; Ferguson et al., 2016), as well as other proteins that participate in endocytosis (Sirotkin et al., 2010; Taylor et al., 2011; Kukulski et al., 2016; Kaksonen et al., 2006; Balzarotti et al., 2017; Gwosch et al., 2020; Schmidt et al., 2021), while electron tomography has been able to resolve membrane shapes during endocytosis (Avinoam et al., 2015; Kukulski et al., 2012). However, neither of the methods can capture both spatial and temporal information at the same time. Under conventional fluorescence microscopy, the clathrin-coated pits appear as diffraction-limited spots due to their small size, which is typically ~ 30 − 150nm in mammals (McMahon and Boucrot, 2011) and yeast (Kukulski et al., 2012), and shape information of the membrane is completely lost (Cocucci et al., 2012; Loerke et al., 2009; Picco et al., 2015; Sirotkin et al., 2010; Kural and Kirchhausen, 2012). On the other hand, super-resolution fluorescence microscopy has been able to reveal the protein organization at the endocytic pit (Mund et al., 2018; Cocucci et al., 2012; Arasada et al., 2018) and to reconstruct the shape of the clathrin coat (Sochacki et al., 2017; Scott et al., 2018; Mund et al., 2023) from averaging over ensembles of endocytic sites. Correlative light and electron microscopy (CLEM) method has exploited the fluorescence of fiducial markers to locate endocytic sites while resolving membrane shapes using electron tomography. However, both super-resolution and CLEM requires sample fixation, therefore, one can identify multiple endocytic sites at the same time and perform the average, yet unable to trace a single endocytic site over time. The temporal information is nevertheless lost.

As a result of the incomplete information obtained by existing experimental methods, both hypotheses have experimental support. Experiments that apply electron microscopy to resolve the membrane shapes of endocytic pits favor the constant area model (Avinoam et al., 2015; Bucher et al., 2018; Sochacki and Taraska, 2019; Sochacki et al., 2021). However, super-resolution imaging combined with analysis of the fluorescence intensity of the clathrin coat is inclined towards the constant curvature model (Willy et al., 2021). In addition, it was argued that the energetic cost of bond reorganization in a regular hexagonal lattice in the constant area model may be too large to be fulfilled (Frey et al., 2020; Frey and Schwarz, 2020; Kirchhausen et al., 2014).

Extensive theoretical efforts have been dedicated to model membrane morphology during endocytosis (Fu and Johnson, 2023), of which molecular dynamics simulations (Varga et al., 2020) and continuum mechanics (Hassinger et al., 2017; Ma and Berro, 2021; Walani et al., 2015; Agrawal and Steigmann, 2008; Rangamani et al., 2013) are two common approaches. Hybrid models were also broadly applied to gain higher resolution than continuum mechanics and lower computing expense than molecular dynamics (Fu et al., 2019, 2021). However, most theoretical investigations have focused on how mechanical properties, such as membrane tension and bending rigidity of the clathrin coat, influence the membrane morphology. The process of curvature generation is either neglected or taken for granted. Only few of them have addressed the difference between the constant area model and the constant curvature model Mund et al. (2023).

In fact, the constant curvature and constant area models are only two extreme models for clathrin assembly during endocytosis and any change in area or curvature are possible at any time point during endocytosis. In this paper, we extend the classic Helfrich theory for membrane deformation to incorporate energy terms associated with clathrin assembly and curvature generation, and compare geometric features calculated by theory with those extracted from experimental data. The negative gradient of the total free energy defines a pathway that neither fits the constant area model nor the constant curvature model. We find that a pathway that is close to the constant area model fits electron tomograms of the endocytic pits the best. Our study also offers experimental suggestions to distinguish between the two main hypotheses.

Models and methods

We model the membrane patch of the CCP (clathrin-coated pit) as a surface which is rotationally symmetric with respect to the z-axis. The shape of the membrane is parameterized with the meridional curve {r(s), z(s)}, where s denotes the arc length along the curve. The bending energy of the membrane (together with the clathrin coat) assumes the Helfrich model (Helfrich, 1973) where κ denotes the bending rigidity of the CCP, C1 and C2 denote the principal curvatures of the surface, C0 denotes the intrinsic curvature of the membrane induced by the clathrin coat. To model a finite area of the clathrin coat, we assume the intrinsic curvature C0 spatially varies as where a denotes positions on the membrane. Here, we choose a to be the surface area calculated from the tip of the membrane, and C0 equals c0 for area a < a0, and rapidly drops to zero when a > a0. The parameter α controls the sharpness of the drop. In the constant area model, we vary the intrinsic curvature c0 but keep the coating area a0 constant, while in the constant curvature model, we vary the coating area a0 but keep the intrinsic curvature c0 constant. As a result of the clathrin coat, the bending rigidity κ also varies as where κcoat and κbare denote the bending rigidity of the clathrin-coated membrane and bare membrane, respectively. Besides the bending energy, the membrane tension contributes to the free energy in the form of where σe denotes the membrane tension at the base and A denotes the surface area of the membrane patch within a fixed radius of Rb. The membrane tension σe and bending rigidity κ define a characteristic length (Derényi et al., 2002). The total free energy Etot = Eb + Et is a functional of the membrane shape. We emphasize that the constant area model refers to the assumption that the clathrin-coated area a0 remains fixed. Meanwhile, the membrane tension σe at the base is held constant, allowing the total membrane area A to vary in response to deformations induced by the clathrin coat. Given a coating area a0 and an intrinsic curvature c0, we numerically solve the variational equations of the energy functional to obtain membrane shapes that minimizes Etot. This approach is based on the assumption that the membrane remains in mechanical equilibrium throughout the process of clathrin assembly, which is justified by the disparity between the timescales of membrane relaxation and clathrin assembly. The characteristic relaxation time of a lipid membrane is given by , where µ ≈ 5 × 10−9 N ⋅ s ⋅ m−1 is the membrane viscosity Arroyo and DeSimone (2009), R0 ≈ 50 nm is the vesicle size Avinoam et al. (2015), and κ ≈ 20 kBT is the bending rigidity Yuan et al. (2021). Substituting these values yields a relaxation time of τ ≈ 1.5 × 10−4 s, which is several orders of magnitude shorter than the clathrin assembly timescale which is approximately one minute. Therefore, it is reasonable to assume that the membrane rapidly equilibrates and remains in mechanical equilibrium during the assembly process. More detailed descriptions of the model can be found in Appendix 1.

We rescale lengths by the characteristic length L0 and energies by 2πκbare. Dimensionless quantities are denoted with a bar. The dimensionless tension energy, defined as Ēt = Et/(2πκ), simplifies to Ēt = Ā/2, where Ā = A/(2πL2) is the dimensionless membrane area. Therefore, the dimensionless membrane tension becomes a constant 1/2. When presenting data, dimensionless quantities are shown on the left axes, and the corresponding dimensional values are shown on the right axes.

Results

Difference between the constant area model and the constant curvature model in terms of membrane morphology

Vesiculation requires assembly of a clathrin coat on the membrane, as well as curvature generation from the clathrin coat. A vesiculation process defines a pathway in the phase space (a0, c0) of the clathrin coat area a0 and the intrinsic curvature c0 of the coat. The constant area model and the constant curvature model are pathways that are made of a vertical line and a horizontal line. Besides these two extreme cases, there is a continuum spectrum of pathways with simultaneously increasing coating area a0 and intrinsic curvature c0 that could lead to vesiculation. Along the pathway, the membrane evolves from a flat shape to a dimple shape, and finally to an Ω-shape, as shown in Figure 2. Hereafter we use the maximal tangential angle ψmax of the membrane as an indicator of the progression of vesiculation - when the membrane is flat, ψmax = 0°, and when the membrane becomes spherical, ψmax = 180°. In our simulation, the neck becomes extremely narrow before ψmax = 180°. Therefore, we consider vesiculation occurs when ψmax reaches 150° (Figure 2a).

Evolution of membrane morphology and phase diagram of vesiculation in the coating area vs. intrinsic curvature (a0, c0) parameter space.

(a) Membrane shapes at different stages of invagination and definition of some variables used in this paper. We define the distance from the axisymmetric axis to the edge of the coating area as Rcoat, the radius of the tangential curvature circle at the tip of the shape as Rt, and the distance from axisymmetric axis to the boundary as Rb. The maximum tangential angle of the cross section contour is ψmax. (b,c) Tip radius Rt vs. maximal angle ψmax for the constant area model in (b) and for the constant curvature model in (c). Dotted lines in (b) denote the analytical solutions. Insets show the ratio of the tip radii Rt at ψmax = 150° and ψmax = 30°. The inset dark dots denotes the numerical results and the red line is the analytical solution (See Appendix 5). (d,e) Coat radius Rcoat vs. maximal angle ψmax for the constant area model in (d) and for the constant curvature model in (e). Dotted lines in (d) denote the analytical solutions. Insets show the ratio of the coat radius Rcoat at ψmax = 150° and ψmax = 90°. The inset dark dots denotes the numerical results and the red lines are the analytical ones (See Appendix 5). (f) Vesiculation diagram in the phase space of (a0, c0). Each horizontal line represents a path of the constant curvature model and each vertical line represents a path of the constant area model. Each path terminates when ψmax = 150°. The solid grey lines represent contours of ψmax = 30°, 60°, 90°, 120°, 150°, respectively. The solid black line is the analytical results for the vesiculation line (See Appendix 4). The dashed black line is a random-picked straight line connecting the origin and the vesiculation boundary. The intersections of the dashed black line and the gray lines are plotted in orange dots and they are the coordinates of (a0, c0) where the five shapes in (a) are located. Shapes are arranged in an increasing order of ψmax in (a). (b-e) Parameters with a bar over them (left vertical axes) are normalized to be dimensionless, and the dimensional parameters (right vertical axes) are calculated by one of the typical fitting values L0 = 40nm (Figure 5)

First, we analyze the difference between the two models in terms of membrane morphology evolution along their pathways. We fit a circle around the membrane tip and use the radius Rt of the circle to characterize the curvature of the membrane at the tip. When Rt is plotted against the maximal angle ψmax, both models show that Rt decreases with increasing ψmax (Figure 2b and c). We stress that even though the intrinsic curvature c0 is fixed in the constant curvature model, it does not imply the tip radius along the vesiculation pathway is a constant. When the coating area is small, the geometric curvature at the membrane tip remains small and differs from the intrinsic curvature. Tip radius in the constant curvature model decays more steeply with ψmax than in the constant area model. This difference becomes obvious when one plots the ratio of the tip radius at ψmax = 150° and ψmax = 30° (Figure 2b and c insets). For the constant area model, the ratio approximately equals to a constant 0.268 regardless of the coating area a0, which agrees well with the analytical result (See Appendix 5). For the constant-curvature model, the ratio remains close to 1 only at small values of c0, as expected from the schematic representation of the model in Figure 1. However, as c0 increases, the deviation from this idealized picture becomes increasingly pronounced.

Another difference between the two models is the evolution of the projected area of the clathrin coat on the substrate. We use the maximal radius Rcoat of the membrane within the clathrin-coated area as the indicator of the projected area (Figure 2a). In the constant area model, Rcoat decreases with increasing ψmax, while in the constant curvature model, Rcoat increases with ψmax and reaches a plateau (Figure 2d and e). The ratio Rcoat (150°)/Rcoat (90°) is about 0.732 in the constant area model and around 1 in the constant curvature model (Figure 2d and e, insets). The analytical curves of Rt and Rcoat against ψmax also fit perfectly with numerical solutions in the constant area model (Figure 2b and d, compare dotted and solid curves). Our calculations therefore demonstrate that the two models exhibit clear differences in the evolution of Rt and Rcoat at the beginning of endocytosis (i.e. when ψmax is small) which can be determined from shapes of endocytic pits obtained experimentally.

To demonstrate how the coating area a0 and the intrinsic curvature c0 of the clathrin coat influence the membrane morphology, for each pair of (a0, c0), we calculate the corresponding membrane shapes and plot the contour lines for ψmax which indicate the stage of endocytosis (Figure 2f). The contour line with ψmax = 150° represents the critical line where vesiculation occurs. The line can be well fitted by the analytical expression (Figure 2f, thick black line, see Appendix 4). It implies that a small intrinsic curvature of the clathrin coat is able to induce vesiculation of the membrane with a large clathrin coat. We find that the distances between contour lines for higher values of ψmax is smaller than those for lower values of ψmax. It means that at the late stage of endocytosis, a small change in a0 and c0 could result in a more dramatic change in the membrane shape than that at the early stage. This trend is demonstrated clearly from the orange dots in Figure 2f, which correspond to shapes in Figure 2a.

Note that in order to produce the phase diagram (Figure 2f), it requires that the bending rigidity of the clathrin coated area κcoat is significantly larger than κbare in the uncoated area. If κcoat is comparable with κbare, there exists a region in the phase diagram in which a single (a0, c0) corresponds to three possible membrane shapes (See Appendix 2—Figure 1a-d). Physically, it implies a discontinuous transition in the membrane shape along a path that passes through this region, and a gap in the maximal angle ψmax would appear. Because in experiments, a wide spectrum of ψmax are observed and no gap in ψmax is found (Avinoam et al., 2015), we keep κcoat much greater than κbare for the rest of the paper. In this regime, the membrane shapes evolve continuously along any pathway that connects the origin (0, 0) with a point on the critical vesiculation curve.

Vesiculation needs free energy sources to drive clathrin assembly and curvature generation

In the previous section, we take curvature generation in the constant area model and clathrin assembly in the constant curvature model for granted, so that the coating area a0 and the intrinsic curvature c0 are imposed, such that the physical forces behind clathrin assembly and curvature generation are ignored. However, the bending energy Eb and the tension energy Et typically increase with a0 and c0 along a vesiculation pathway. Therefore, vesiculation will be energetically unfavorable if no additional free energy sources are provided. In this section, we extend the model to include free energy terms for the assembly of the clathrin coat and its reorganization for curvature generation.

To describe the assembly of the clathrin coat in the constant curvature model, we introduce where µ denotes the effective surface binding energy density of clathrin molecules with the membrane. This term reduces the free energy with increasing a0, therefore, driving the assembly of clathrin. We can identify three types of free energy curves for different assembly strength µ: (1) When µ is small, the total free energy Etot = Eb + Et + Ea as a function of the coating area a0 has two local minima, with the lowest one at a small a0 and the other one at the maximum a0 where vesiculation occurs (Figure 3a, red curve). The minima are separated by an energy barrier that is significantly higher than the thermal energy kBT and the clathrin coat would assemble to a small area and halt. (2) With increasing µ, the lowest free energy minimum is shifted to the vesiculation point, but the energy barrier still exists and the clathrin coat remains small (Figure 3a, orange curve). (3) Vesiculation could happen for large enough µ such that the energy barrier vanishes and the free energy Etot monotonically decreases with a0 (Figure 3a, green). Based on the above analysis of the energy landscape we construct the phase diagram of the constant curvature model with clathrin assembly in the phase space of (c0, µ) and classify the points into four types. Besides the three types mentioned above, when the intrinsic curvature c0 is small, increasing a0 to its maximum value (105nm2) cannot produce vesiculation. (Figure 3b, gray region). The critical assembly energy density µ at which the energy barrier vanishes is found to increase with the intrinsic curvature c0 (Figure 3b, interface between the green region and the orange region), which implies that a larger assembly strength of clathrin coat µ is needed to complete vesiculation if the clathrin coat has a higher intrinsic curvature c0. When comparing the contour lines of the energy barrier ΔEtot = 1kBT and ΔEtot = 10kBT (Figure 3b, dotted curve and dash-dotted curve), the gap between them increases with c0, which means that the energy efficiency is reduced with c0 in the sense that, for larger c0, a larger increase in µ is needed to reduce the same amount of free energy.

Free energy evolution in the constant curvature and constant area models when accounting for one of either the polymerization energy term Ea = −µa0 or the curvature generation energy term (type 1) and (type 2).

(a,b) Free energy landscape of the modified constant curvature model where with polymerization energy Ea = −µa0 in (a) and the corresponding phase diagram in the phase space of (c0, µ) in (b). The rightmost endpoint of each curve is the vesiculation point where ψmax = 150°. The red line in (a) and red dots in (b) correspond to pathways with minimum free energy Etot appearing at a point other than the vesiculation point on the Etot − ā0 curve. The orange line in (a) and the orange dots in (b) correpsond to pathways where the vesiculation point is the minimum free energy point, but an energy barrier still exists. The green line in (a) and green dots in (b) correspond to vesiculation pathways without an energy barrier. The energy barrier ΔEtot is defined as the energy difference between the maximum point and the first minimum point before the maximum. ΔEtot of the orange curve is shown in (a) as a typical example. The gray dots in the left-hand side of (b) correspond to pathways that numerically fail to reach the vesiculation point when ā0 reaches its upper limit 10. (c,d) Free energy landscape of the modified constant area model where ā0 = 2 with curvature generation energy Ec = −νa0c0 in (c) as a function of the intrinsic curvature c0 and the corresponding phase diagram in the phase space of (a0, ν) in (d). The gray dots in the left-hand side of (d) correspond to pathways that numerically fail to reach the vesiculation point when reaches its upper limit 5. (e,f) Free energy landscape of the modified constant area model where ā0 = 2 with in (e) and the corresponding phase diagram in the phase space of (a0, ν) in (f). The gray dots in the left-hand side of (f) correspond to pathways that numerically fail to reach the vesiculation point when reaches its upper limit 5. (a-f) The parameters with a bar over them are normalized to be dimensionless, and the dimensional parameters are calculated using one of the typical fitting values L0 = 40nm (Figure 5). (b,d,f) The black line separates the region wiht gray dots (which did not numerically reach vesicultion) from the other regions. The dotted, dash-dotted and solid gray lines respectively represent the paths where ΔEtot = 1kBT, ΔEtot = 10kBT, ΔEtot = 100kBT.

We next consider curvature generation in the constant area model. As the molecular mechanisms of curvature generation of the clathrin coat remains debated, we introduce a phenomenological model in which the free energy has the general form, where ν denotes the strength of curvature generation, m and n are two positive numbers that are associated with the molecular mechanisms of curvature generation. The free energy Ec in Equation 6 decreases with increasing c0, therefore, driving curvature generation. We set m = 1 such that Ec is proportional to the coating area. Note that m cannot be zero, otherwise, Ec only depends on the intrinsic curvature c0 and can be nonzero even when the coating area is zero. As for the power n of the intrinsic curvature c0, we set n = 1 or 2 (called Model(1,1) and Model(1,2), respectively). Physically, Model(1,2) implies cooperativity in the curvature generation such that the reduction of free energy per increase of unit curvature is proportional to the current curvature, i.e., ΔEc ∝ −c0Δc0, while in Model(1,1), the reduction of free energy per increase of unit curvature is independent of current curvature. For Model(1,1), when ν is small, the total free energy Etot = Eb + Et + Ec as a function of the intrinsic curvature c0 has two minima, the lowest one at a small positive c0 and the other one at the maximum c0 where vesiculation occurs (Figure 3c, red curve). Further curvature generation is strictly limited by the high energy barrier (sometimes more than 100kBT) between the two minima. With increasing ν, the lowest minimum shifts to the vesiculation point, but the energy barrier still prevents curvature generation (Figure 3c, orange line). For a large enough ν, the free energy monotonically decreases with c0 and the curvature generation proceeds until vesiculation occurs (Figure 3c, green curve). When the coating area a0 is very small, vesiculation fails to occur even when the intrinsic curvature is increased to its maximum value (0.125nm−1) (Figure 3c, gray region). In the phase space of (a0, ν), the critical value of ν where the energy barrier vanishes increases with the coating area a0 (Figure 3d, interface between the orange region and the green region), which implies that a larger clathrin coat needs a stronger strength of curvature generation to complete vesiculation.

Model(1,2) has similar free energy landscape as Model(1,1) (Compare Figure 3c and e, d and f). However, in Model(1,2), for very small ν, the lowest free energy minimum is strictly pinned at c0 = 0, which implies no spontaneous curvature generation. In contrast, the minimum is at a small positive c0 in Model(1,1), which indicates slight curvature generation.

Determination of the vesiculation pathway from the energy landscape

In this section, we combine the assembly energy Equation 5 and the curvature generation energy Equation 6 together and calculate the total free energy Etot (a0, c0) = Eb +Et +Ea +Ec as a function of both the coating area a0 and the intrinsic curvature c0. A pathway from the origin can be constructed by the descent along the negative gradient of the free energy landscape −∇Etot. In Figure 4a we show the free energy landscape for a fixed assembly strength and varied reorganization strength for model(1,2). When ν is small, the energy contour lines near the origin are kinked, which represents an energy barrier that prevents the path from going up, i.e. from generating curvature. The path extends horizontally and terminates on the a0 − axis (Figure 4a, first column, red curve). With increasing ν, the kinked contour lines shift towards larger a0 and the path can be lifted up to in the middle and drops to the a0 − axis in the end (Figure 4a, second column, orange curve). Beyond a critical ν, the energy barrier vanishes and the path bends up and terminates on the vesiculation curve (Figure 4a, third column, green curve). Further increasing µ leads to the path lifting up at a smaller coating area a0 (Figure 4a, fourth column, green curve). The path goes horizontally first and is later lifted up, which resembles the path of the constant area model. The three types of pathways are classified into three colored regions in the phase diagram of (µ, ν), which represent complete vesiculation (green), partial vesiculation (orange) and no vesiculation (red), respectively (Figure 4c).

Vesiculation phase diagram when accounting for both the polymerization energy µ and the reorganization energy ν.

(a,b) Free energy landscape for Model(1,2) (i.e. with reorganization energy ) in (a) and Model(1,1) (i.e. with reorganization energy ) in (b). Thick black lines are the analytical solutions for the vesiculation boundary. Thin rainbow-colored lines visually represent the energy landscape (values of the color bar are for the dimensionless free energy scale). Thick colored-lines represent pathways that stream along the negative gradient of the free energy landscape in the phase space starting from the origin (i.e. no clathrin assembled and no curvatuve). Our model shows that only a subset of suitable (µ, ν) values create pathways that lead to vesiculation, i.e. that reach the thick black line (thick green lines in the third and fourth panels). The orange and red curves are pathways that fail to reach vesiculation. The red curve does not produce any curvature, while the orange curve generates a small curvature but never leads to vesiculation. (c,d) Phase diagrams for Model(1,2) in (c) and Model(1,1) in (d) show the relationship between pathway types and the (µ, ν) values. The colors of the dots correspond to the same types of pathways as represented by thick colored lines in (a,b). Our results show that larger µ or ν values lead to an easier vesiculation. Parameters with a bar over them are normalized to be dimensionless, and the dimensional parameters are calculated using one of the typical fitting values L0 = 40nm (Figure 5).

The free energy landscapes of Model(1,1) dramatically differs from Model(1,2). When ν is small, the energy gradient is strongly biased towards the horizontal direction, and the path extends horizontally with little or no curvature generation (Figure 4b, first column). For an intermediate ν, the path first goes towards the top right direction until and then slowly bends down and extends towards large coating area along a valley formed in the energy landscape, which corresponds to membrane shapes with a small dimple (Figure 4b, second column). For large enough ν, the path shoots nearly straightly towards the top right direction before it reaches the vesiculation line (Figure 4b, third column). Further increasing ν makes the path more straight and terminates at a smaller coating area (Figure 4b, fourth column). The (µ, ν) phase diagram shows the parameter regions that lead to complete, partial or no vesiculation for Model(1,1) (Figure 4d).

Comparison between different models with the experimental data

The constant area model and the constant curvature model represent two extreme pathways of membrane vesiculation. We have found constant-area-like pathways in Model(1,2) and straight-line-like pathways in Model(1,1). In order to understand which model is the most plausible, we compare membrane shapes predicted by the models with the membrane profiles obtained by electron microscopy in (Avinoam et al., 2015). The fitting error ϵ of a vesiculation path reflects the relative difference between the model-predicted geometric features along the path and the rolling median of the corresponding experimental data (Figure 5 and Appendix 3). The fitting geometric features include neck width, tip radius, and invagination depth (Figure 5c). We draw the corresponding optimum energy paths that minimize the fitting error (Figure 5b), and compare the best model-predicted shapes with the experimental ones (Figure 5d).

Comparison between our theory and experimental data from mammalian cells.

(a) Parameter fit of the best of the four models (constant curvature model, constant area model, Model(1,1), Model(1,2)) to obtain the minimum error ϵ. Fitting procedure of Model(1,1) and Model(1,2) consider the total free energy Etot = Eb + Et + Ea + Ec, while the fitting of the constant area model and the constant curvature model consider Etot = Eb + Et. The optimized parameters are ā0 ∈ [0, 10] for the constant area model, for the constant curvature model, and for model(m,n), and L0 ∈ [10nm, 100nm] within an interval of 10nm in the four models. We only assign fitting errora to the parameter sets that lead to vesiculation and only plot the error figure for the best L0. (b) Vesiculation pathways with minimum fitting error in the four models. In each model, we use the best L0 value from (a) to obtain the dimensional scale of the (a0, c0) phase space. (c) Comparison of model fits and experimental data for three geometric features: neck width, tip radius (Rt) and invagination depth. Neck width is calculated as the distance between the left and right parts of the shape for ψmax = 90°, and the invagination depth is measured as the height from the base to the tip of the invagination. (d) Comparison between the model-predicted shapes and the experimental shapes. Experimental membrane shapes for mammalian cells are grouped according to their maximum angle as a proxy for the different stages of CME. The number of experimental shapes falling in a certain ψmax range is defined as n. The black lines are the average experimental shapes after symmetrization. The model-predicted shapes are calculated by the midpoint value of each ψmax interval. (c,d) The curves predicted by theory are shown with colored lines, and experimental data is shown with gray dots and black lines. Parameters with a bar over them are non-dimensionalized. The detailed procedure to treat the experimental data can be found in Appendix 3.

For Model(1,2) and Model(1,1), the fitting parameters include the polymerization strength and the reorganization strength which together determine the vesiculation pathway, as well as the characteristic length L0 which scales the size of the membrane. For Model(1,2), the best fits are obtained for and L0 = 30nm (Figure 5a). The resulting path moves horizontally at first and then bents up vertically (Figure 5b), which resembles the behavior of the constant area model. For Model(1,1), the best fitting parameters are and L0 = 40nm (Figure 5a). The fitting path is almost a straight line towards the vesiculation line (Figure 5b). The optimum fitting error of Model(1,2) (ϵ = 0.14) is slightly better than that of Model(1,1) (ϵ = 0.17).

We also perform the fitting procedure to the constant area model and find the optimum parameters are a0 = 1.69 × 104nm2 and L0 = 30nm. For the constant curvature model, the best fitting parameters are c0 = 0.043nm−1 and L0 = 50nm. The minimum fitting error of the constant curvature model (ϵ = 0.28) is exactly twice as large as that of the constant area model (ϵ = 0.14) (Figure 5a and b). So considering the fitting error and the pathway in (a0, c0) phase diagram, we raise the conclusion that the experimental vesiculation process probably favors constant-area-like pathways.

When comparing the model-predicted geometric features with the rolling median of the experimental data, we find that the four models fit almost equally well the experimental data for the neck width(Figure 5c left). Model(1,2) and the constant area model predict very similar results such that the curves almost overlap with each other (Figure 5c, red curve and orange curve). The predictions of these two models match the rolling median of the experimental data best. The constant curvature model strongly deviates from the rolling median of the experimental tip radius, particularly in the early stage of vesiculation when ψmax < 90°.

To facilitate comparison between the axisymmetric membrane shapes predicted by the model and the non-axisymmetric profiles obtained from electron microscopy, we apply a symmetrization procedure to the experimental data, which consist of one-dimensional membrane profiles extracted from cross-sectional views, as detailed in Appendix 3 (see also Fig. 1). Then, we average the symmetrized profile within an interval of ψmax ∈ [ψ0 −10°, ψ0 +10°] and overlay the averaged profile with model-predicted shapes for ψmax = ψ0 (Figure 5d). At the early stage when the membrane exhibits a dimple shape (0° < ψmax < 60°), the membrane morphology predicted by the constant curvature model is distinguishable from the other three models, particularly when looking at the tip radius. At the late stage when the membrane exhibits an Ω-shape, i.e., ψmax > 90°, the difference in shape between models is mainly manifested in the invagination depth. The constant curvature model and Model(1,1) mainly predict a deeper invagination depth than the symmetrized experimental profile, while the constant area model and Model(1,2) usually give much better fitting.

Discussion

Three types of clathrin coats

In this paper, we have constructed a physical model to describe how curvature generation and clathrin assembly are interrelated during the vesiculation process in CME. Previous experiments have reported three groups of clathrin coated pits, which are plaques, abortive pits, and pits that lead to vesiculation, according to their structural and dynamic properties (Maupin and Pollard, 1983; Ehrlich et al., 2004; Loerke et al., 2009; Saffarian and Kirchhausen, 2009; Kirchhausen, 2009; Lampe et al., 2016).

In Figure 4, we show that depending on the clathrin assembly strength µ and its reorganization strength ν, the pathway might end up with three possible final shapes: (i) a flat membrane with no curvature generation, (ii) a nearly flat membrane with small curvature generation, (iii) a spherical cap that leads to vesiculation. They essentially correspond to the three types of clathrin-coated pits found in experiments. Based on the phase diagram of Model(1,2) (Figure 4c), the difference between the three groups comes from the difference in the assembly and reorganization strengths of clathrin molecules. Furthermore, Model(1, 2) predicts that at the boundary between the type (iii) region and the type (ii) region, the reorganization strength ν increases with the assembly strength. This result has important implications to explain why large plaques are commonly observed in experiments. The large area of the plaques are due to the strong assembly strength µ. However, for these plaques to go to vesiculation, a strong reorganization energy ν is also needed. Therefore, the combination of a strong µ and weak ν leads to the formation of large plaques. Model(1,2) predicts that a plaque or an abortive pit can be transformed into a vesicle by either increasing the reorganization strength or reducing the assembly strength (Figure 6). The former ends up with a large vesicle and the latter ends up with a small vesicle. This can be used as a test of our theory with experiments to modify the binding affinity of clathrin molecules with adaptor proteins on the membrane. Weakening the affinity might increase the portion of vesicles and reduce the portion of plaques, though the vesicles would become smaller.

Tip radius of vesiculaion shapes (Rves) in Model(1,2).

The colored region shows the (µ, ν) sets that lead to vesiculation, and brighter colors correspond to larger Rves. Decreasing assembly strength µ or increasing reorganization strength ν might lead to vesiculation of different vesicle sizes. An example from (µ, ν) = (13.3 × 10−3kBT ⋅ nm−2, 8.8kBT) to the vesiculation region is marked by arrows, red dots and corresponding vesicle shapes. The characteristic length L0 = 30nm is used in the calculation.

Cooperativity in the curvature generation process

In Figure 5, we show the best fitting results for all the four models and find that Model(1,2) produces better fits than Model(1,1), which suggests the existence of cooperativity in the curvature generation process. In particular, if curvature generation is driven by breaking bonds in the hexagonal lattice, cooperativity implies that the number of newly broken bonds is proportional to the number of already broken bonds. Because of this cooperativity, at the early stage of endocytosis bonds are broken slowly and clathrin assembly dominates over curvature generation. At the late stage of endocytosis, an increasing number of bonds are broken and curvature generation could happen rapidly and dominate over clathrin assembly. Altogether, this cooperativity leads to a constant-area-like behavior. Similar effect have been reported in Ref. (Mund et al., 2023).

Furthermore, although our model is purely energetic and does not explicitly incorporate dynamics, we observe in Figure 4(a) that along the green curve—representing the trajectory predicted by model (1,2)—the total free energy Etot exhibits a much sharper decrease at the late stage (near the vesiculation line) compared to the early stage (near the origin). This suggests a transition from slow to fast dynamics during endocytosis. Such a transition is consistent with experimental observations, where significantly fewer number of images with large ψmax are captured compared to those with small ψmax (Mund et al., 2023).

Saturation effect at high membrane curvatures

Note that our model involves two distinct concepts of curvature growth. The first is the growth of imposed curvature—referred to here as intrinsic curvature and denoted by the parameter c0—which is driven by the reorganization of bonds between clathrin molecules within the coat. The second is the growth of the actual membrane curvature, reflected by the increasing value of ψmax. The latter process is driven by the former.

Models (1,1) and (1,2) incorporate energy terms (Equation 6) that promote the increase of intrinsic curvature c0, which in turn drives the membrane to adopt a more curved shape (increasing ψmax). In the absence of these energy contributions, the system faces an energy barrier separating a weakly curved membrane state (low ψmax) from a highly curved state (high ψmax). This barrier can be observed, for example, in the red curves of Figure 3(a–c) and in Appendix 6–Figure 1. As a result, membrane bending cannot proceed spontaneously and requires additional energy input from clathrin assembly.

The energy terms described in Equation 6 serve to eliminate this energy barrier by lowering the energy difference between the uphill and downhill regions of the energy landscape. However, these same terms also steepen the downhill slope, which may lead to overly aggressive curvature growth. To mitigate this effect, one could introduce a saturation-like energy term of the form where cs represents a saturation curvature. Importantly, adding such a term would not alter the conclusions of our study, since the energy landscape already favors high membrane curvature (i.e., it is downward sloping) even without the additional energy terms.

The difference in membrane morphology between the different models is most salient at the early stage of endocytosis

When we compare the model predictions, we find that the difference in membrane morphology between models is not as big as expected, which might explain why it has been difficult to distinguish between the constant area and the constant curvature models for so long. For instance, the neck width vs. ψmax and the invagination depth vs. ψmax are similar for all the four models (Figure 5c, left and right). The best fitting error of the four models calculated from the geometric features are relatively close, except for the constant curvature model, which gives the worst fit (Figure 5a). The models are mainly distinguishable from the tip radius vs. ψmax plot at the early stage of endocytosis when the membrane is nearly flat (Figure 5c, middle), i.e. for shapes with small ψmax. However, published experimental shape at early stages of endocytosis are sparse. Our result hints that in order to distinguish between the models, collecting membrane shapes at the early stage is necessary and the relation of tip radius vs. ψmax is the key geometric feature to tell the models apart.

The projected area of clathrin coat in the plane of the plasma membrane could distinguish between the two models

In Figure 2d and e we have shown that the coat radius Rcoat, which represents the projected area of the clathrin coat in the plane of the plasma membrane, as a function of ψmax exhibit opposite trends for the constant curvature model and the constant area model. The results suggest that in experiments the projected area for the constant area model would first increase and then decrease over time, finally reaching a plateau. However, for the constant curvature model, the projected area would increase over time and finally reach a plateau without a decreasing phase. This result suggests another method to distinguish between the two models via the projected area measurement. The idea has been used in a study where the clathrin-coated pit was imaged with platinum replica and cryoelectron microscopy and tomography (Sochacki et al., 2021). The results support a constant-area-like model, consistent with the prediction of our Model(1,2), in which the dome structures have a slightly larger coating area than flat structures. On the other hand, another study has used the super-resolved live cell fluorescence imaging with TIRF to measure the growth of the clathrin coat over time (Willy et al., 2021). The authors found a smooth drop in the projected area of clathrin coat over time. However, based on a computer simulation of clathrin assembly, they concluded that the smooth drop of the projected area was the result of a constant-curvature-like model because a constant-area-like model would produce a sharp drop. We attribute the difference between their model and our model to the fact that they model the clathrin coat as a discrete lattice while we use a continuum mechanics method. More importantly, in their model, the moment at which curvature generation occurs was arbitrarily imposed at 80% of clathrin triskelions. If the transition were chosen to occur with fewer triskelions, e.g. 40%, the sharp drop in the project area might not happen in the constant-area-like model. Furthermore, the authors used the number of triskelions to monitor the progress of endocytosis which terminates when the triskelions reach the maximum number. This choice might bias towards the constant-curvature-like model because the vesiculation may not happen at all when the clathrin numbers reaches its maximum.

The bending rigidity of the coated area should be much larger than the uncoated area

Comparison of our model to experimental data demonstrates that the relative bending rigidities of the coat and the membrane are constrained. Indeed, if κcoat /κbare < 50, the model predicts an abrupt change (or a gap) in ψmax at the end of vesiculation (See Appendix 2—Figure 1), i.e., a snap-through transition (Hassinger et al., 2017). If a gap in ψmax existed, we would expect that the distribution of experimental shapes to be discontinuous, with no or very few data corresponding to a certain range of ψ. However, in the experiments (Avinoam et al., 2015), the endocytic pits shapes are continuously distributed across the ψmax spectrum, indicating the ergodicity of ψmax during the endocytic process. Our calculation suggests that the clathrin coat is about 50 times stiffer than the membrane.

Comparison with particle wrapping

The purpose of the clathrin-mediated endocytosis studied in our work is the recycling of membrane and membrane-protein, and the cellular uptake of small molecules from the environment — molecules that are sufficiently small to bind to the membrane or be encapsulated within a vesicle. In contrast, the uptake of larger particles typically involves membrane wrapping driven by adhesion between the membrane and the particle, a process that has also been studied previously (Góźdź, 2007; Bahrami et al., 2016). In our model, membrane bending is driven by clathrin assembly, which induces curvature. In particle wrapping, by comparison, the driving force is the adhesion between the membrane and a rigid particle. In the absence of adhesion, wrapping increases both bending and tension energies, creating an energy barrier that separates the flat membrane state from the fully wrapped state. This barrier can hinder complete wrapping, resulting in partial or no engulfment of the particle. Only when the adhesion energy is sufficiently strong can the process proceed to full wrapping. In this context, adhesion plays a role analogous to curvature generation in our model, as both serve to overcome the energy barrier. If the particle is spherical, it imposes a constant-curvature pathway during wrapping. However, the role of clathrin molecules in this process remains unclear and will be the subject of future investigation.